---
title: "Final Project-Telco Customer Churn"
author: "brandon rumbarger"
date: "2026-04-29"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(caret)
library(pROC)
library(ggplot2)
```

```{r}
data <- read.csv("C:/Users/brumb/OneDrive/Documents/R/WA_Fn-UseC_-Telco-Customer-Churn.csv")


str(data)
summary(data)

```

**Data Preparation**
```{r}
data <- data %>%
  mutate(
    TotalCharges = as.numeric(ifelse(TotalCharges == "", NA, TotalCharges)),
    Churn = as.factor(Churn)
  ) %>%
  na.omit()

data <- na.omit(data)

data$Churn <- as.factor(data$Churn)
```

**Train-Test Split**
```{r}
set.seed(123)

trainIndex <- createDataPartition(data$Churn, p = 0.7, list = FALSE)
train <- data[trainIndex, ]
test  <- data[-trainIndex, ]


train$customerID <- NULL
test$customerID  <- NULL
```

**Model Development**
```{r}
model <- glm(Churn ~ tenure + MonthlyCharges + Contract,
             data = train,
             family = "binomial")

summary(model)

```

**Model Evaluation**
```{r}
probabilities <- predict(model, test, type = "response")

predictions <- ifelse(probabilities > 0.5, "Yes", "No")
predictions <- as.factor(predictions)
```

```{r}
confusionMatrix(predictions, test$Churn)
```
**Roc Curve**
```{r}
roc_obj <- roc(test$Churn, probabilities)

plot(roc_obj, col = "blue", main = "ROC Curve")
auc(roc_obj)

```

**History of Monthly Changes**
```{r}
ggplot(data, aes(x = MonthlyCharges, fill = Churn)) +
  geom_histogram(bins = 30, alpha = 0.6)
```
Customers with higher monthly charges are more likely to churn, as seen by the greater concentration of churned customers at higher price levels.

**Contact Type vs Churn**
```{r}
ggplot(data, aes(x = Contract, fill = Churn)) +
  geom_bar(position = "fill") +
  ylab("Proportion") +
  ggtitle("Churn Rate by Contract Type")
```
Customers on month-to-month contracts have significantly higher churn rates compared to those on long-term contracts.

**Odds Ratios**
```{r}
exp(coef(model))
```
Higher monthly charges increase churn risk
Longer tenure reduces churn
Long-term contracts significantly reduce churn
